Building a continuous benchmarking ecosystem in bioinformatics

### 2024-10-30 manifesto-main.tex
Ben's

Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific. In this case, they can also be seen as (dataset-specific) metadata. To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.

@@

Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods). One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a new method. Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation. Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then harder to establish. Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial benchmark formalization could even provide the basis for the formal preregistration. In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains difficult. Other benchmarking models are possible, from challenges to hackathons, with the hope that a consensus of the masses can further help to establish neutrality. One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template. Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}.

@@

\begin{itemize}
    \item Flexibility vs. complexity: unrestricted freedom for method contributors can mean displacing the burden of maintainability towards the benchmarker. Fewer restrictions (e.g., simpler entry points) means that more methods can be included in a single benchmark; but, it can also imply a higher complexity for the initial setup by the benchmarker. Constraints can also enable validation, which can greatly help with the development process and quality assurance.
    \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}. A benchmarker can state the desired degree of replicability and constraints; different tasks require different levels of compliance. 
    \item  Security concerns: decentralized runs lower the participation barrier, but also increase the attack  surface. Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigation. Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
\end{itemize}

>>>>

Dataset-specific parameters, e.g., the true number of clusters when evaluating clustering algorithms, are not method-specific. They can also be seen as (dataset-specific) metadata. To differentiate between the effect of the parameter and the methods performances themselves, data-specific parameters should be shared across methods during evaluation.

@@

Neutrality is a highly desirable property in a benchmark: it means that efforts have been taken to ensure impartiality or unbiasedness (or at least transparency) of the various choices made (e.g., what datasets to include, how exactly to run methods, how to evaluate methods)\cite{weber2019-el,jelizarow2010-kr}. One could claim that all MDP benchmarks are non-neutral, since the goal is generally to highlight the virtues of a new method. Similarly, ground truth generation and simulations in MDP benchmarks can unconsciously mirror the method capabilities, mechanisms, or strengths, thus biasing evaluation. Furthermore, there are many field experts who may wish to conduct a benchmark (or participate in a collaborative benchmark) on analyses that involve their own tools, where neutrality is then harder to establish. Here,  recent work to mitigate this conflict involves pre-registration and posting the parameters of the study in advance at a registry \cite{Sullivan2019-wh, Olevska2021-if}; initial benchmark formalization could even provide the basis for the formal preregistration. In contrast, BOP benchmarks give, in principle, a higher degree of neutrality, but the challenge to assess neutrality directly remains difficult. Other benchmarking models are possible, from challenges to hackathons, with the hope that a consensus of the masses can further help to establish neutrality. One potential strategy, in the context of the systematic design of benchmarks, is establishing a system where a benchmark definition is parsed into a pre-registration template. Templates exist for simulation studies that could be adapted more generally for benchmarking projects \cite{Siepe2023-tz}.

@@

\begin{itemize}
    \item Flexibility vs. constraints:  An unconstrained benchmark grants method contributors greater flexibility but requires more effort from the benchmark maintainer (maintenance, version support). Introducing constraints (such as software dependencies or common data formats) adds initial setup work but can boost maintainability through built-in validation and quality checks.
    \item  Achieving software reproducibility needs discipline and commitment \cite{Lamb2021-tb}. A benchmarker can state the desired degree of replicability and constraints; different tasks require different levels of compliance. 
    \item  Security concerns: decentralized runs lower the participation barrier, but also increase the attack  surface. Sandboxed environments (restricting network access, filesystem access etc) can be provided as mitigation. Additionally, static and dynamic code analysis in the continuous integration process can aid in approving potentially risky submissions.
\end{itemize}
